Galois Characterization of Function Fields of \\ Transcendence Degree $1$ over Number Fields

### 2017-01-27 Master Thesis/synopsis.tex
Fix typos

\begin{proof}
The compositum $K_{v_1}K_{v_2}$ in a fixed separable closure of $K$ is henselian with respect to both $v_1$ and $v_2$, hence separably closed by Theorem \ref{theorem-fkschmidt}. Therefore:
\[1=G_{K_{v_1}K_{v_2}} = G_{K_{v_1}}\cap G_{K_{v_2}} = G_{v_1}\cap G_{v_2} \qedhere\]
\end{proof}

@@

\paragraph{} Let $n\in\mathbb{N}$. In the previous remark, we have determined that the following diagram commutes: 
\[ \begin{tikzcd}
G_K \ar[r]\ar[d, "\sigma"'] & G_k\\
G_{K'}\ar[ru]
\end{tikzcd} \]
Now the operation of $G_K$ and $G_{K'}$ on $\mu_n$ factors through $G_k$, thus we have $\sigma(g)\xi = g\xi$ for all $\xi\in\mu_n$ and $g\in G_K$. Therefore, for all intermediate fields $L$ of $\overline{K}/K$, $\sigma$ induces an isomorphism:
\[ \varphi_n:L'^\times/n \cong\mathrm{H}^1(L', \mu_n) \to \mathrm{H}^1(L,\mu_n) \cong L^\times/n \]
Obviously, for intermediate fields $L\subset M$ of $\overline{K}/K$ the following diagram commutes:
\[ \begin{tikzcd}
L'^\times/n \ar[r, "\varphi_n"]\dar & L^\times/n \dar\\
M'^\times/n \ar[r, "\varphi_n"'] & M^\times/n
\end{tikzcd} \]
Let $P\in X_0$ and $v_P:K\to\mathbb{Z}\cup\{\infty\}$ be the corresponding discrete normalized valuation and let $U_P$ be the group of units of the valuation ring $\mathcal{O}_P$. We denote the natural projection $U_P\to k(P)^\times,\ x\mapsto x\mod P$ by $r_P$.

@@

\begin{proof}
Observe that $\varphi(-1)^2 = \varphi(1) = 1$, and since $\varphi$ is bijective on $K'^\times$, it follows $\varphi(-1) \neq 1$ and hence $\varphi(-1) = -1$. Now, to prove the assertion, it suffices to show $\varphi(x + y) = \varphi(x) + \varphi(y)$ for all $x,y\in K'^\times$.
\begin{itemize}
\item \textit{Case 1:} If $x + y = 0$, we have:
\[ \varphi(x + y) = 0 = \varphi(x) + \varphi(-1)\varphi(x) = \varphi(x) + \varphi(-x) = \varphi(x) + \varphi(y) \]
\item \textit{Case 2:} Let $x + y\neq 0$ and $S\subset X_0$ be a finite subset such that $\varphi(x + y)$, $\varphi(x)$ and $\varphi(y)$ are contained in $K^S$. Looking at the commutative diagram in \mbox{Step 1} we get
\begin{align*}
r_P(\varphi(x+ y)) &= \varphi_P(r_{P'}(x + y)) \\
&= \varphi_P(r_{P'}(x)) + \varphi_P(r_{P'}(y)) = r_P(\varphi(x) + \varphi(y))
\end{align*}
for all $P\in X_0\setminus S$. Note that $\varphi_P: k(P')\to k(P)$ is a field isomorphism and that the residue maps $r_{P}: \mathcal{O}_{P} \to k(P)$ are ring homomorphisms. Since $K^S\to \prod_{P\in X_0\setminus S} k(P)^\times$ is injective, we get $\varphi(x+y) = \varphi(x) + \varphi(y)$.\qedhere
\end{itemize}
\end{proof}

>>>>

\begin{proof}
Let $K_{v_i}$ be a henselization of $K$ with respect to $v_i$. The compositum $K_{v_1}K_{v_2}$ in a fixed separable closure of $K$ is henselian with respect to both $v_1$ and $v_2$, hence separably closed by Theorem \ref{theorem-fkschmidt}. Therefore:
\[1=G_{K_{v_1}K_{v_2}} = G_{K_{v_1}}\cap G_{K_{v_2}} = G_{v_1}\cap G_{v_2} \qedhere\]
\end{proof}

@@

\paragraph{} Let $n\in\mathbb{N}$. In the previous remark, we have reduced the problem, so that the following diagram commutes: 
\[ \begin{tikzcd}
G_K \ar[r]\ar[d, "\sigma"'] & G_k\\
G_{K'}\ar[ru]
\end{tikzcd} \]
Now the operation of $G_K$ and $G_{K'}$ on $\mu_n$ factors through $G_k$, thus we have $\sigma(g)\xi = g\xi$ for all $\xi\in\mu_n$ and $g\in G_K$. Therefore, for all intermediate fields $L$ of $\overline{K}/K$, $\sigma$ induces an isomorphism:
\[ \varphi_n:L'^\times/n \cong\mathrm{H}^1(L', \mu_n) \to \mathrm{H}^1(L,\mu_n) \cong L^\times/n \]
Obviously, for intermediate fields $L\subset M$ of $\overline{K}/K$ the following diagram commutes:
\[ \begin{tikzcd}
L'^\times/n \ar[r, "\varphi_n"]\dar & L^\times/n \dar\\
M'^\times/n \ar[r, "\varphi_n"'] & M^\times/n
\end{tikzcd} \]
Let $P\in X_0$ and $v_P:K\to\mathbb{Z}\cup\{\infty\}$ be the corresponding normalized discrete valuation and let $U_P$ be the group of units of the valuation ring $\mathcal{O}_P$. We denote the natural projection $U_P\to k(P)^\times,\ x\mapsto x\mod P$ by $r_P$.

@@

\begin{proof}
Observe that $\varphi(-1)^2 = \varphi(1) = 1$, and since $\varphi$ is bijective on $K'^\times$, it follows $\varphi(-1) \neq 1$ and hence $\varphi(-1) = -1$. Now, to prove the assertion, it suffices to show $\varphi(x + y) = \varphi(x) + \varphi(y)$ for all $x,y\in K'^\times$.
\begin{itemize}
\item \textit{Case 1:} If $x + y = 0$, we have:
\[ \varphi(x + y) = 0 = \varphi(x) + \varphi(-1)\varphi(x) = \varphi(x) + \varphi(-x) = \varphi(x) + \varphi(y) \]
\item \textit{Case 2:} Let $x + y\neq 0$ and $S\subset X_0$ be an admissible finite subset such that $\varphi(x + y)$, $\varphi(x)$ and $\varphi(y)$ are contained in $K^S$. Looking at the commutative diagram in \mbox{Step 1} we get
\begin{align*}
r_P(\varphi(x+ y)) &= \varphi_P(r_{P'}(x + y)) \\
&= \varphi_P(r_{P'}(x)) + \varphi_P(r_{P'}(y)) = r_P(\varphi(x) + \varphi(y))
\end{align*}
for all $P\in X_0\setminus S$. Note that $\varphi_P: k(P')\to k(P)$ is a field isomorphism and that the residue maps $r_{P'}: \mathcal{O}_{P'} \to k(P')$ are ring homomorphisms. Since $K^S\to \prod_{P\in X_0\setminus S} k(P)^\times$ is injective, we get $\varphi(x+y) = \varphi(x) + \varphi(y)$.\qedhere
\end{itemize}
\end{proof}
